---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 


```{r}
library(dplyr)
library(tidyr)
library(stringr)

av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)

parse_number <- function(x) as.numeric(str_extract(x, "[0-9]+"))

# reshaping the data for deaths
deaths <- av %>% select(URL, starts_with("Death")) %>% pivot_longer(cols = Death1:Death5,
                                                                           names_to = "death",
                                                                           values_to = "result") %>% group_by(URL) %>% summarise(death_count = sum(result == "YES"))  %>% summarise(death_mean = mean(death_count))

print(deaths$mean_n)


```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

### Include the code

### Include your answer

Vedant - "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team. Let’s put it this way: If you fall from four or five stories up, there’s a 50 percent chance you die. Getting a membership card in the Avengers is roughly like jumping off a four-story building."

```{r}

# filter avengers who died at least once

av_deaths <- av %>%
  filter(Death1 == "YES" | Death2 == "YES" | Death3 == "YES" | Death4 == "YES" | Death5 == "YES")

num_deaths <- nrow(av_deaths)

percentage_deaths <- (num_deaths / nrow(av)) * 100

percentage_deaths

```

If they joined the avengers team, there is a 40% chance that the character died. I did this by calculating the number of total death and divide by total avangers which gives us 39.88 or approx 40%. Hence we have verified the fact.


Abhi - There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife


Upload your changes to the repository. Discuss and refine answers as a team.
